---
title: "RprojectScript"
output:
  html_document: default
  pdf_document: default
date: "2023-07-31"
---

## Setup libraries and folders

```{r setup, include=FALSE}
# Remove  all objects from R memory 
rm(list=ls())  

# Set a path to the library containing packages
.libPaths("C:/Computerteknologi/10_semester/Rcourse")
.libPaths()


# Load selected packages# 
pack<-c("car","sandwich","lmtest","RColorBrewer","mgcv","foreign","xtable"
        ,"AER","stargazer", "MASS")

lapply(pack, require, character.only=T)

# Check the working path
getwd()

# Specify  paths
getwd()

# Set working path
data_path<-"C:/Computerteknologi/10_semester/Rcourse/project/data"
data_graph <-"C:/Computerteknologi/10_semester/Rcourse/project/graph"
data_results <-"C:/Computerteknologi/10_semester/Rcourse/project/result"
data_scripts <-"C:/Computerteknologi/10_semester/Rcourse/project/script"
setwd(data_path)

#Check the directory
dir()

data = read.csv("mortality2.csv",header = FALSE,skip = 0)


```

## Preprocessing of data

```{r}

data_countries = data[!grepl("exclud",data$V1),]
data_countries = data[!grepl("income",data$V1),]

data_income = data[grepl("income",data$V1),]
data2 = data_countries[-(1:3),-1]



rownames(data2) = data_countries[-(1:3),1]
colnames(data2) = data_countries[3,-(1:1)] 


data2 = data2[,-4:-33]
data2 = data2[,-35:-37]
data2 = data2[rowSums(is.na(data2))<5,]

subset(data2,)

#data2 = data[,-4:-33]

```

Webscraping webpage to get areas, unions and continents out

```{r}
#Loading the rvest package
library('rvest')

#Specifying the url for desired website to be scraped
url <- 'https://www.iban.com/country-codes'


#Reading the HTML code from the website
webpage <- read_html(url)

drivers_F1 <- html_element(webpage, "table.table") %>%
  html_table()

sum1 = sum(data2$`Country Code` %in% drivers_F1$`Alpha-3 code`)
#webpage %>% html_elements("table#Country-Codes-A-C")
#drivers_F1$Country
data3=data2[!data2$`Country Code` %in% drivers_F1$`Alpha-3 code`,]
data3=data3[data3$`Country Code`!="XKX",]

data_categories=data3

```

HDI list:

```{r}
#Loading the rvest package
library('rvest')

#Specifying the url for desired website to be scraped
url <- 'https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index'


#Reading the HTML code from the website
webpage <- read_html(url)

drivers_F1 <- html_element(webpage, "table.sortable") %>%
  html_table()

hdi_list = drivers_F1[,c("Rank","Nation")]
hdi_list = hdi_list[-1,]
#webpage %>% html_elements("table#Country-Codes-A-C")
#drivers_F1$Country
#data3=data2[!data2$`Country Code` %in% drivers_F1$`Alpha-3 code`,]
#data3=data3[data3$`Country Code`!="XKX",]

#data_categories=data3
```

## Analytics of data

-   Connection between immortality and HDI

-   Developing over time

-   Countries from lowest to highest immortality

-    Area of countries

-   Comparison of Low / medium / high income countries with

plots:

-   Map with the locations of the world countries'
